---
title: "MATH1324 Assignment 1"
subtitle: Modeling Body Measurements
output:
  html_notebook: default
---

## Student Details


Jeremy Jian Zhi Chia (s3917515)


## Problem Statement

Write a clear and concise problem statement that guides your investigation. Explain which variable you choose and outline the approaches taken for normal distribution fitting.

The biiliac width refers to the largest measurement of the pelvis between the outside borders of the upper iliac bones in human, more commonly known as the pelvic bone width. The variable chosen in this investigation is the biiliac diameter. In general, the distribution of biiliac diameter among adults are similar and their measurement values are close to the mean, with a lower frequency of higher/lower measurement values. That being said, values near to the mean are rather common, while those further from the mean are less common.
- discuss about general distribution of biiliac width

The purpose of this investigation is to determine if the biiliac diameter in men and women fit a normal distribution.

## Load Packages

```{r}
# Import dependencies
library(readr)
library(dplyr)
library(tidyr)
```

## Data

Import the body measurements data and prepare it for analysis. Show your code.


```{r}
# Import dataset
bdims <- read_csv("bdims-1.csv")

# View structure of dataset
str(bdims)

# Set gender as factor
bdims$sex <- factor(bdims$sex, levels = c(0, 1), labels = c("female", "male"))

# Check attributes of gender factor
attributes(bdims$sex)
```
```{r}
# Selecting only variables of interest from entire dataset
biiliac_gender <- bdims %>% dplyr::select(bii.di, sex)

# View selected variables
biiliac_gender %>% head()
```


## Summary Statistics


Calculate descriptive statistics (i.e., mean, median, standard deviation, first and third quartile, interquartile range, minimum and maximum values) of the selected measurement grouped by sex.

```{r}
# Create new variables and filter by gender
biiliac_male <- biiliac_gender %>% filter(sex == "male")
biiliac_female <- biiliac_gender %>% filter(sex == "female")
```

```{r}
# Mean of male biiliac width
male_mean <- biiliac_male$bii.di %>% mean()
male_mean
```
```{r}
# Mean of female biiliac width
female_mean <- biiliac_female$bii.di %>% mean()
female_mean
```
```{r}
# Median of male biiliac width
male_median <- biiliac_male$bii.di %>% median()
male_median
```
```{r}
# Median of female biiliac width
female_median <- biiliac_female$bii.di %>% median()
female_median
```
```{r}
# SD of male biiliac width
male_sd <- biiliac_male$bii.di %>% sd()
male_sd
```
```{r}
# SD of female biiliac width
female_sd <- biiliac_female$bii.di %>% sd()
female_sd
```

```{r}
# Quartiles of male biiliac width
male_quantile <- biiliac_male$bii.di %>% quantile()
male_quantile
```
The first quartile for males biiliac width is 26.8 and the third quartile is 29.45.

```{r}
# Quartiles of male biiliac width
female_quantile <- biiliac_female$bii.di %>% quantile()
female_quantile
```
The first quartile for females biiliac width is 26.2 and the third quartile is 29.2.

```{r}
# IQR of male biiliac width
male_iqr <- biiliac_male$bii.di %>% IQR()
male_iqr
```
```{r}
# IQR of female biiliac width
female_iqr <- biiliac_female$bii.di %>% IQR()
female_iqr
```
```{r}
# Min/Max of male biiliac width
male_range <- biiliac_male$bii.di %>% range()
male_range
```
```{r}
# Min/Max of male biiliac width
female_range <- biiliac_female$bii.di %>% range()
female_range
```

## Distribution Fitting

Compare the empirical distribution of selected body measurement to a normal distribution separately in men and in women. You need to do this visually by plotting the histogram with normal distribution overlay. Show your code.


```{r}
# Plotting distribution for males billiac width
hist(biiliac_male$bii.di, freq = FALSE, breaks = 12, xlab = "Biiliac Diameter (cm)", main = "Histogram of Male Biiliac Diameter (cm)")

# Adding empirical density line
lines(density(biiliac_male$bii.di), col = "red")

# Adding normal density line
lines(seq(male_range[1], male_range[2], by=.5), dnorm(seq(male_range[1], male_range[2], by=.5),
      male_mean, male_sd), col="blue")

# Adding legends
legend(19, 0.20, legend=c("Empirical Density", "Normal Density"), col=c("red", "blue"), lty=1:1, cex=0.8)
```
```{r}
# Plotting distribution for female billiac width
hist(biiliac_female$bii.di, freq = FALSE, breaks = 12, xlab = "Biiliac Diameter (cm)", main = "Histogram of Female Biiliac Diameter (cm)")

# Adding empirical density line
lines(density(biiliac_female$bii.di), col = "red")

# Adding normal density line
lines(seq(female_range[1], female_range[2], by=.5), dnorm(seq(female_range[1], female_range[2], by=.5),
      female_mean, female_sd), col="blue")

# Adding legends
legend(19, 0.20, legend=c("Empirical Density", "Normal Density"), col=c("red", "blue"), lty=1:1, cex=0.8)
```
## Interpretation

Going back to your problem statement, what insight has been gained from the investigation? Discuss the extent to how your theoretical normal distribution fits the empirical data.


